The Concurrent Calculi Formalisation Benchmark

### 2024-02-22 short-paper/short-paper.tex
Fix some minor issues.

In this spirit, we introduce a new collection of benchmark problems
specifically crafted to tackle common challenges encountered during
the mechanisation of formal models for concurrent and distributed
programming languages.
We particularly focus on process calculi as these provide a simple but
realistic showcase of common issues encountered during mechanisation
efforts.
Concurrent calculi are notably subtle: for instance, it took nine
years before the incorrect proof of type safety in the seminal paper
by Honda et al.~\cite{Honda1998} was rectified in Yoshida and
Vasconcelos’s subsequent work\cite{Yoshida2007}.\todo{Marco: rewrite this to be correct}
Similarly, several key results in papers on session types have turned out
to be false later on~\cite{Gay2020,10.1145/3290343}, illustrating the
value of machine-checked proofs.\todo{Rephrase this sentence to be more clear on the references}

@@

Given those peculiarities, the syntax and semantics of $\pi$-calculi have been
mechanised from an early age (see~\cite{Melham1994}) with many proof
assistants and in so many encoding styles that it would take too long to
review.
\todo{However, there is no standard/consensus on what to do... elaborate}
They range from concrete encodings based on named syntax
\cite{Melham1994} to basic de Bruijn~\cite{Hirschkoff1997, Perera2018} and locally nameless
representation %~\cite{Chargueraud2012}
(\eg~\cite{Castro2020}).
Nominal approaches are also common (see~\cite{Bengtson2009}), but they
may be problematic in proof assistants based on dependent types.
Another approach is shallow embeddings~\cite{Castro-Perez2021}, which avoids the need to explicitly represent names.
The case study \cite{AmbalLS21} compares four approaches to encoding
binders in Coq as a first step towards developing tools for working
with higher-order process calculi.  They found that working directly
with de Bruijn indices was easiest and shortest since the existing
approaches developed for $\lambda$-calculus binders do not work well
with scope extrusion.

>>>>

In this spirit, we introduce a new collection of benchmark problems
specifically crafted to tackle common challenges encountered during
the mechanisation of formal models for concurrent and distributed
programming languages.
We particularly focus on process calculi as these provide a simple but
realistic showcase of common issues encountered during mechanisation
efforts.
Concurrent calculi are notably subtle: for instance, it took nine
years before the incorrect proof of type safety in the seminal paper
by Honda et al.~\cite{Honda1998} was rectified in Yoshida and
Vasconcelos’s subsequent work\cite{Yoshida2007}.\todo{Marco: rewrite this to be correct}
Similarly, several key results in papers on session types have turned out
to be false later on (as described in \eg \cite{Gay2020,10.1145/3290343}), illustrating the potential
value of machine-checked proofs.

@@

Given those peculiarities, the syntax and semantics of $\pi$-calculi have been
mechanised from an early age (see~\cite{Melham1994}) with many proof
assistants and in so many encoding styles that it would take too long to
review.
Despite this, almost all of these mechanisations still rely on ad-hoc solutions to accommodate scope extrusion.
They range from concrete encodings based on named syntax
\cite{Melham1994} to basic de Bruijn~\cite{Hirschkoff1997, Perera2018} and locally nameless
representation %~\cite{Chargueraud2012}
(\eg~\cite{Castro2020}).
Nominal approaches are also common (see~\cite{Bengtson2009}), but they
may be problematic in proof assistants based on dependent types.
Another approach is shallow embeddings~\cite{Castro-Perez2021}, which avoids the need to explicitly represent names.\todo{Merge the previous sentence into the HOAS section?}
The case study \cite{AmbalLS21} compares four approaches to encoding
binders in Coq as a first step towards developing tools for working
with higher-order process calculi.  They found that working directly
with de Bruijn indices was easiest and shortest since the existing
approaches developed for $\lambda$-calculus binders did not work well
with scope extrusion.
